Developmental consistency in children's drawings of object categories

### 2021-11-09 writing/Manuscript_June2021.tex
Updates to ms post MCF edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Consistency and variability in the development of children's drawings of object categories},
  pdflang={en-EN},
  pdfkeywords={children's drawings, visual production, tracing, object recognition, visuomotor control},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Consistency and variability in the development of children's drawings of object categories}

@@

\abstract{
Childen's drawings of common object categories become dramatically more recognziable across development. What are the major factors that explain this developmental change? Here, we examined the degree to which these developmental changes in recognizibility vary across different drawing tasks (i.e.~drawing from observation vs.~from memory), geographical locations (San Jose, US vs.~Beijing, China),and with children's tracing abilities. To do so, we collecting digital drawings of object categories (e.g., cat, airplane) from 4-9 year-olds (N=253). Overall, we show broad consistency in these developmental trajectories acros both drawing tasks and these two geographical locations, and we find that children's tracings abilites are good predictors of the recognizability of the drawings that they produce. In addition, we find that children recruited near Beijing, China producedmore recognizable drawings but showed similar tracing abilities to childen recruited near San Jose, CA. Overall, this work suggests that the developmental trajectory of children's drawings are remarkably consistent and not easily explainable by changes in their visuomotor control.
}

@@

In modern times, drawings are produced prolifically by children from an early age.
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities (Minsky \& Papert, 1972), and accordingly a long history of work has examined changes in children's drawings across development (Fury, Carlson, \& Sroufe, 1997; Karmiloff-Smith, 1990; Kellogg, 1969; Piaget, 1929).
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age; younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between target concepts (e.g., \textit{adult} vs.~\textit{child}) than older children, who enrich their drawings with more diagnostic part (Sitton \& Light, 1992) and relational (Light \& Simmons, 1983) information.
Overall, younger children produce drawings of object categories that are less diagnostic of the categories they are trying to depict (Long, Fan, Chai, \& Frank, 2021; Long, Fan, \& Frank, 2018).

What drives these dramatic changes in children's drawings across development?
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements (Freeman, 1987; Rehrig \& Stromswold, 2018).
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints, such as how well children are able to access previously acquired semantic knowledge about each target concept and maintain this information in working memory during drawing production.
Thus, drawing is a skill that likely also relies on children's evolving perceptual category representations (Dekker, Mareschal, Sereno, \& Johnson, 2011; Long et al., 2018; Natu et al., 2016) as well as their increasing working memory capacity (Pailian, Libertus, Feigenson, \& Halberda, 2016).

One idea is that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant perceptual features of different categories: that is, when asked to \enquote{draw a {[}rabbit{]}}, they may struggle to conjure up the relevant perceptual details and then hold in mind what rabbits tend to look like.
On this account, providing children with additional perceptual information about different categories -- for example, canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults (Fan et al., 2020).
However, prior work also suggests that younger children tend to draw what they know about objects rather than integrate the information in their immediate perceptual experience.\\
For example, when asked to draw from observation, younger children tend to include features that are not visible from their vantage point, yet are diagnostic of category membership (e.g., a handle on a \textit{mug}) (Barrett \& Light, 1976; Bremner \& Moore, 1984), and only omit these features later in development.
Similarly, young children will insist that their nearly identical drawings of different concepts (e.g., balloon and person) unambiguously refer to different things (Bloom \& Markson, 1998).
Thus, an alternative possibility is that only older children may be able to produce more recognizable drawings when provided with canonical exemplars of different categories.
On this account, changes in children's drawings from memory across age may be largely due to other factors beyond changes in memory -- for example, changes in how children represent the diagnostic visual features of each category (Long et al., 2018).

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when asked to draw from memory (\enquote{Can you draw a {[}rabbit{]}?) and when asked to draw a canonical exemplar from observation (}Can you draw this {[}rabbit{]} as it looks in the picture?") across two different geographical sites (San Jose, USA and Beijing, China).
To do so, we gathered digital drawings from a large sample of children 4-9 years of age in a controlled experimental setting.
While children in different communities may spend more or less time practicing drawing or use different visual conventions to draw (Cohn, 2012; Willats, 2006), most empirical studies on children's drawings have been conducted exclusively on children from the United States -- or focused on differences in educational practices between communities (Huntsinger, Jose, Krieg, \& Luo, 2011; La Voy et al., 2001; Winner, 1989).
In addition, we assessed the degree to which visuomotor development accounts for developmental changes in drawing recognizability by measuring each child's shape tracing abilities.

We had several predictions about the nature of the effects that we would observe. First, we predicted that only older children would be able to use the visual information present in the canonical photographs to improve their drawings, in keeping with accounts of naïve realism and contra a strong account of working memory limitations.
Second, we predicted that we would see convergence in the development of drawing abilities across both geographical sites, with older children becoming progressively better at producing recognizable drawings.
Finally, we predicted that most of the variance across geographical sites in drawing ability would be explainable by differences in visuomotor control, operationalized as performance on a shape tracing task; these primary analyses were pre-registered at \url{https://osf.io/qymjr/}.

@@

265 participants were recruited from the San Jose Children's Discovery museum, the Palo Alto Junior Museum and Zoo, and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e.~20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose based sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of 121 children. In the Beijing based sample, 132 children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed 11.46 to analysis (min 6, max = 12). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at {[}blinded{]} (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at {[}blinded{]} in Beijing, China.

@@

Before beginning, a trained experimenter first told each child \enquote{After this game is over, someone is going to try to recognize what you were trying to draw. So, please draw so that someone else could try what you were trying to draw.} A native English speaker gave these instructions to the San Jose based sample, and a native Mandarin speaker gave a translation of these instructions to the Beijing based sample. The rest of the instructions were given via audio and video cues of native English/Mandarin speakers, respectively; all experimental code, videos, and stimuli are available on the public repository for this project. All children used their fingers to draw on an Ipad Pro (12.9") that was set at a slight angle on a table via a tablet case. All experimental code, videos, and data are available on the public repository for this project.

Children first completed two series of tracing trials to obtain an estimate of their tracing abilities; children were first asked to trace a square and then a complex shape (see Figure \ref{fig:example-tasks}). Children had maximum of 30 seconds to produce their tracings/drawings on all trials, and strokes could not be deleted once drawn; stroke-by-stroke data for each drawing was stored on a remote server.

The primary experimental manipulation was the drawing task that children participated in; each child completed both tasks and the order of tasks was counterbalanced across participants. To elicit drawings from memory, children saw a short video clip where an experimenter said, \enquote{What about a {[}cat{]}? Can you draw a {[}cat{]}?} (exact Mandarin translation has trouble in latex). To elicit drawings from observation, children heard audio of the same experimenter who said, \enquote{What about this {[}cat{]}? Can you draw the {[}cat{]} as it looks in the picture?} (Mandarin translation) while seeing a canonical photograph of each category; this photograph then remained on the screen for the duration of the drawing trial. This photograph was randomly sampled from one of three possible exemplars. Overall, each child completed 2 tracing trials and 12 drawing trials, with six categories randomly assigned to each condition and displayed in a random order. Task examples are shown in Figure \ref{fig:example-tasks}.

@@

As in (Long et al., 2018), we used an automated procedure for evaluating how accurately participants performed the tracing task, validated against empirical judgments of tracing quality. We decomposed tracing accuracy into two components: a shape error component and a spatial error component. Shape error reflects how closely the participant's tracing matched the contours of the target shape; the spatial error reflects how closely the location, size, and orientation of the participant's tracing matched the target shape.

To compute these error components, we applied an image registration algorithm, AirLab (Sandkühler, Jud, Andermatt, \& Cattin, 2018), to align each tracing to the target shape, yielding an affine transformation matrix that minimized the pixel-wise correlation distance between the aligned tracing, \(T\), and the target shape, \(S\): \(Loss_{NCC} = - \frac{\sum S \cdot T - \sum E(S) E(T)}{N \sum Var(S) Var(T)}\), where \(N\) is the number of pixels in both images. The shape error was defined by the final correlation distance between the aligned tracing and the target shape. The spatial error was defined by the magnitude of three distinct error terms: location, orientation, and size error, derived by decomposing the affine transformation matrix above into translation, rotation, and scaling components, respectively. In sum, this procedure yielded four error values for each tracing: one value representing the shape error (i.e., the pixel-wise correlation distance) and three values representing the spatial error (i.e., magnitude of translation, rotation, scaling components).

We used the same tracing quality ratings to obtained in (Long et al., 2018) to assign weights to each of their error terms; adult observers (\(N\)=70) rated 1325 tracings (i.e., 50-80 tracings per shape per age) and evaluated \enquote{how well the tracing matches the target shape and is aligned to the position of the target shape} on a 5-point scale. An ordinal regression mixed-effects model to predict these 5-point ratings, which contained correlation distance, translation, rotation, scaling, and shape identity (square vs.~star) as predictors, with random intercepts for rater. This model yielded parameter estimates that could then be used to score each tracing in the dataset; we averaged scores for both shapes within session to yield a single tracing score for each participant.

@@

We assessed the recognizability of each drawing via an online recognition experiment. Adult participants based in the U.S. were recruited via Prolific for a 15 minute experiment,compensated at \$14/hour, and recognized a random subset of around 140 drawings that were balanced with respect to age, category, and site. Each drawing was recognized by 10 participants. Participants were shown a were shown these drawings in a random order and asked \enquote{What does this look like?}. Participants selected an answer from the twelve categories that children were prompted to draw and were were encouraged to guess. No participants were excluded from analysis for missing a catch trial (a free response describing the task that were completing). These binary recognition scores were then averaged across participants to yield a recognition score (e.g., recognized by 80\% of participants) for each drawing.

@@

We used a combination of deep CNN activations and logistic regressions to obtain automated recognition scores, as per our pre-registered protocol. To encode the high-level visual features of each sketch, we used the VGG-19 architecture \cite{simonyan2014very}, a deep convolutional neural network pre-trained on Imagenet classification. We used model activations in the second-to-last layer of this network, which is the first fully connected layer of the network (FC6). Raw feature representations in this layer consist of flat 4096-dimensional vectors, to which we applied channel-wise normalization across all filtered drawings in the dataset. Next, we used these features to train object category decoders. We then trained a 12-way logistic classifier with L2 regularization (tolerance = .1, regularization = .1), and used this classifier to estimate the category label for each drawing in the dataset.To avoid any bias due to imbalance in the distribution of drawings over categories, we randomly under sampled such that there were an equal number of drawings for each combination of geographical site (San Jose, Beijing) and the 12 categories. No additional metadata about the age of the child who produced each sketch was provided to the decoder. This procedure was repeated for each drawing in the dataset, yielding a binary a recognition score for each drawing.

@@

To assess our main hypotheses, we fit linear mixed effects models to the human recogniton scores to assess the factors that influenced the recognizability of the drawings that children produced. A first linear mixed effect model was fit to the recognizability scores assigned to each drawing, including fixed effects of children's age (in years), geographical site (San Jose vs.~Beijing), and drawing task (drawing from perception vs.~memory) and the three way interaction between these key variables. We initially planned to include random slopes for the effect of drawing task on each child (as this varied within-subjects), and random slopes for the effect the full three-way interactions between task, age, and site on each category. However, all models with this complicated random effects structure failed to converge, and the reported models the maximal random effects structure that did converge.

In a secondary analysis, we aimed to understand the degree to which any of the above effects were mediated by (1) children's tracing abilities and (2) the amount of effort that children expended while drawings. We thus ran the same main model while also now including fixed effects of children's estimated tracing score (see above), the time children spent drawing (in seconds), the mean intensity of the drawing (i.e.~percentage of non-white pixels), and the number of strokes children used. All predictors were scaled to have a mean of 0 and a standard deviation of 1. Finally, we also assessed the degree to which tracing ability development differed by geographical site, where tracing scores were modeled as a function of age (in years), site, and their interaction, with random intercepts for each participant.

This anlaysis plan was pre-registered at \url{https://osf.io/qymjr/}; however, we had initially planned to use automated recognition scores as our main dependent variable, following prior work (Long et al., 2021); however, we found that these automated recognition scores were only modestly correlated with human recognition scores (\(r\) = 0.40, \(t\) = 0.40, p \textless{} .001); furthermore, descriptive plots of the automated recognition scores revealed a difference between the two drawing tasks that was not evident in the human data. To be conservative, we thus fit all of our models to the human recognition data but have made all model classifications available for future researcs.

@@

Overall, we found relative consistency in the developmental trajectory of children's drawings. Figure \ref{fig:main-results} shows the recognizability of children's drawings at each age as a function of the drawing task and the geographical site that they were located in; All model coefficients can be found in Table 1. First, we found consistent and steady changes in the recognizability of children's drawings across this age range, replicating prior work with observational datasets collected at a free-standing kiosk (Long et al., 2021). Critically, these changes in recognizability did not depend on the drawing task children completed: children's drawings were equally recognizable when they were asked to \enquote{Draw this rabbit as it looks in the picture} and they were simply asked \enquote{What about a rabbit? Can you draw a rabbit.} While we anticipated that older children might produce more recognizable drawings when drawing from observation -- as do adults when presented with canonical photographs of object categories (Yang \& Fan, 2021) -- this was not the case.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Developmental consistency in children's drawings of object categories},
  pdflang={en-EN},
  pdfkeywords={children's drawings, visual production, tracing, object recognition, visuomotor control},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\title{Developmental consistency in children's drawings of object categories}

@@

\abstract{
Childen's drawings of common object categories become dramatically more recognziable across development. What are the major factors that explain this developmental change? Here, we examined the degree to which these developmental changes in recognizibility vary across different drawing tasks (i.e.~drawing from observation vs.~from memory), geographical locations (San Jose, US vs.~Beijing, China), and with children's tracing abilities. To do so, we collecting digital drawings of object categories (e.g., cat, airplane) from 4-9 year-olds (N=253). Overall, we show broad consistency in these developmental trajectories acros both drawing tasks and these two geographical locations, and we find that children's tracings abilites are good predictors of the recognizability of the drawings that they produce. In addition, we find that children from Beijing, China produced more recognizable drawings but showed similar tracing abilities to childen from San Jose, USA. Overall, this work suggests that the developmental trajectory of children's drawings are remarkably consistent and not easily explainable by changes in their visuomotor control or working memory capacity.
}

@@

In modern times, drawings are produced prolifically by children from an early age.
Figurative drawings have long provided inspiration for scientists investigating children's emerging cognitive abilities (Minsky \& Papert, 1972), and accordingly a long history of work has examined changes in children's drawings across development (Fury, Carlson, \& Sroufe, 1997; Karmiloff-Smith, 1990; Kellogg, 1969; Piaget, 1929).
Indeed, there appear to be dramatic changes in how children encode diagnostic visual information in their drawings across age; younger children (4-5 years) tend to include fewer cues in their drawings to differentiate between target concepts (e.g., \textit{adult} vs.~\textit{child}) than older children, who enrich their drawings with more diagnostic part (Sitton \& Light, 1992) and relational (Light \& Simmons, 1983) information.

What drives these dramatic changes in children's drawings across development?
A common view is that these changes are driven primarily by children's increasing ability to plan and control their motor movements (Freeman, 1987; Rehrig \& Stromswold, 2018).
While such changes in visuomotor control are clearly important, this view fails to account for other important constraints --- including how children represent different visual concepts and how well children are able to access these visual concepts when producing drawings.

In our prior work, we found that evidence that changes in children's drawings partly reflect changes in children's visual representations of common object categories. In a large observational dataset, older children produced drawings of object categories that were more diagnostic of the categories they are trying to depict (Long, Fan, Chai, \& Frank, 2021; Long, Fan, \& Frank, 2018), even when accounting for differences across age in basic shape tracing abilities and the amount of effort children expended on individual drawings. Furthermore, older children tended to rely more on these same diagnostic visual features when recognizing each other's drawings, suggesting that changes in children's internal representations may drive parallel changes in hildren's production and recognition of drawings. Thus, the drawings that children produce likely not only reflect their increasing motor abilities but their evolving perceptual category representations (Dekker, Mareschal, Sereno, \& Johnson, 2011; Natu et al., 2016).

However, this work leaves open the contribution of children's evolving memory: children's ability to access semantic knowledge about categories and to maintain this information in mind when producing a drawing likely changes across childhood (Pailian, Libertus, Feigenson, \& Halberda, 2016).\\
Here, we directly test the idea that a principal reason younger children produce less recognizable drawings is because they simply have more difficulty recalling the relevant perceptual features of different categories: that is, when asked to \enquote{draw a {[}rabbit{]}}, they may struggle to conjure up the relevant perceptual details and then hold in mind what rabbits tend to look like.
On this account, providing children with additional perceptual information about different categories -- for example, canonical photographs of typical exemplars -- could help them improve their drawings of these categories, as it does with adults (Fan et al., 2020).
However, prior work also suggests that younger children tend to draw what they know about objects rather than integrate the information in their immediate perceptual experience.\\
For example, when asked to draw from observation, younger children tend to include features that are not visible from their vantage point, yet are diagnostic of category membership (e.g., a handle on a \textit{mug}) (Barrett \& Light, 1976; Bremner \& Moore, 1984), and only omit these features later in development.
Similarly, young children will insist that their nearly identical drawings of different concepts (e.g., balloon and person) unambiguously refer to different things (Bloom \& Markson, 1998).
Thus, an alternative possibility is that only older children may be able to produce more recognizable drawings when provided with canonical exemplars of different categories.
On this account, changes in children's drawings from memory across age may be largely due to other factors beyond changes in memory -- for example, changes in how children represent the diagnostic visual features of each category (Long et al., 2018).

To tease apart these alternatives, we investigated the development of children's ability to produce recognizable drawings of visual concepts when children were provided with a verbal cue (\enquote{Can you draw a {[}rabbit{]}?) versus when provided with a picture cue (}Can you draw this {[}rabbit{]} as it looks in the picture?\enquote{). On verbal-cue trials, children thus must access their mental representation of a}rabbit" and choose the features necessary to convey that object's identity. Conversely, on picture-cue trials, children are explicitly asked to rely on the visual features provided in a canonical photograph of each object category.

Second, we tested the generalizability of our findings across different populations by recruiting children from two different sites in different countries --- San Jose, USA and Beijing, China.
Most empirical studies on children's drawings have been conducted exclusively on small samples of children from the United States or Western Europe, limiting their generalizability. Further, children in different communities may spend more or less time practicing drawing or use different visual conventions to draw (Cohn, 2012; Willats, 2006). While some prior work has focused on differences in children's drawings across these two countries, most of this descriptive work has focused on differences in educational practices that may account for any observed differences (Huntsinger, Jose, Krieg, \& Luo, 2011; La Voy et al., 2001; Winner, 1989). Instead, here we focus on the degree to which the overall developmental trajectory of drawing abilities are similar between children at these two sites.

Finally, we assessed the degree to which visuomotor development accounts for observed developmental changes in drawing recognizability. While it is clear that visuomotor control influences how and what we can draw, little work has directly related measures of visuomotor control to drawing outcomes {[}(Long et al., 2021). To do so, we measured each child's visuomotor control via a shape tracing task on a table and related these measurements of tracing accuracy to the recognizability of the drawings that each child produced.

In the present study, we thus collect shape tracings and digital drawings of visual concepts from 4-9 year-old children in two different countries using both picture cues and verbal cues. In doing so, we make three major contributions to our understanding of the development of children's drawings. First, we replicate our prior findings (Long et al., 2021) that the recognizability of children's drawings increases steadily throughout this age range (4-9 years). Second, we test the degree to which working memory constraints might account for these developmental changes. In keeping with accounts of naïve realism and contra a strong account of working memory limitations, we predicted that only older children would be able to use the visual information present in the canonical photographs to improve their drawings, Third, we test the generalizability of our findings across two different countries. We predicted that we would see convergence in the development of drawing abilities across both geographical sites, with older children becoming progressively better at producing recognizable drawings. In addition, we predicted that most of the variance across geographical sites in drawing ability would be explainable by differences in visuomotor control, operationalized as performance on a shape tracing task; these primary analyses were pre-registered at \url{https://osf.io/qymjr/}.

@@

265 participants were recruited from two local children's museum in {[}BLINDED{]} and preschool and elementary schools outside of Beijing; approximately equal numbers of participants were recruited in Northern California and the Beijing area. Our goal was to recruit 120 children between 4-9 years of age after exclusions (i.e.~20 4-year-olds, 20 5-year-olds, etc.) at each geographical site. In the San Jose based sample, 135 children participated in the experiment; 6 participants were excluded, (3) for skipping more than 6 drawing trials and (3) for scribbling three or more times in a row. Six additional participants were tested but their data was not recorded due to a technical error, and two participants never advanced past the practice trials, leading to a final sample of 121 children. In the Beijing based sample, 132 children participated; an additional 8 participants were tested but their data was not recorded due to a technical error with the remote database. Two 10-year-olds (aged 10 years, 0 months and 10 years, 1 month) were accidentally tested and included in the 9-year-old age group. On average, each child contributed 11.46 drawings to analysis (min 6, max = 12). No additional demographic data was recorded about the participants. This protocol was approved by both the Institutional Review Board at {[}blinded{]} (43992, Development of Children's Drawing Abilities) and the Department of Psychology Ethics Committee at {[}blinded{]} in Beijing, China.

@@

Children participated in a drawing game using their fingers to draw, completing 2 shape tracing trials and 12 object category drawing trials. Children drew using an Ipad Pro (12.9") that was set at a slight angle on a table via a tablet case, and had maximum of 30 seconds to produce their tracings and drawings. Strokes could not be deleted once drawn, and stroke-by-stroke data for each drawing was stored on a remote server.

Before beginning the game, a trained experimenter first told each child \enquote{After this game is over, someone is going to try to recognize what you were trying to draw. So, please draw so that someone else could try what you were trying to draw.} A native English speaker gave these instructions to the San Jose based sample, and a native Mandarin speaker gave a translation of these instructions to the Beijing based sample.

After these instructions, children first completed two series of tracing trials to obtain an estimate of their tracing abilities; children were first asked to trace a square and then a complex shape (see Figure \ref{fig:example-tasks}). The primary experimental manipulation was the drawing cue that children heard on the subsequent drawing trials (see task examples in Figure \ref{fig:example-tasks}). Children completed six trials with the same cue type (verbal or picture) before switching; cue order was counterbalanced across participants. On verbal cue trials, children saw a short video clip where an experimenter said, \enquote{What about a {[}cat{]}? Can you draw a {[}cat{]}?}. On picture-cue trials, children heard audio of the same experimenter who said, \enquote{What about this {[}cat{]}? Can you draw the {[}cat{]} as it looks in the picture?} while seeing a canonical photograph of each category; this photograph then remained on the screen for the duration of the drawing trial. This photograph was randomly sampled from one of three possible exemplar (see Appendix). All experimental code, videos, translations, and stimuli are available on the public repository for this project.

@@

As in Long et al. (2021), we used an automated procedure for evaluating how accurately participants performed the tracing task, validated against empirical judgments of tracing quality. We decomposed tracing accuracy into two components: a shape error component and a spatial error component. Shape error reflects how closely the participant's tracing matched the contours of the target shape; the spatial error reflects how closely the location, size, and orientation of the participant's tracing matched the target shape. See Appendix for details on how these error components were computed and translated into a tracing score.

@@

We assessed the recognizability of each drawing via an online recognition experiment. Adult participants based in the U.S. were recruited via Prolific for a 15 minute experiment, compensated at \$14/hour, and asked to recognize a random subset of around 140 drawings that were balanced with respect to age, category, and site. Each drawing was recognized by 10 participants. Participants were shown these drawings in a random order and asked \enquote{What does this look like?}. Participants selected an answer from the twelve categories that children were prompted to draw and were encouraged to guess. No participants were excluded from analysis for missing a catch trial (a free response describing the task that were completing). These binary recognition scores were then averaged across participants to yield a recognition score for each drawing.
(e.g., recognized by 80\% of participants).

@@

We used a combination of deep CNN activations and logistic regressions to obtain automated recognition scores, as per our pre-registered protocol. To encode the high-level visual features of each sketch, we used the VGG-19 architecture (Simonyan \& Zisserman, 2014) a deep convolutional neural network pre-trained on Imagenet classification. We used model activations in the second-to-last layer of this network, which is the first fully connected layer of the network (FC6). Raw feature representations in this layer consist of flat 4096-dimensional vectors, to which we applied channel-wise normalization across all filtered drawings in the dataset. Next, we used these features to train object category decoders. We then trained a 12-way logistic classifier with L2 regularization (tolerance = .1, regularization = .1), and used this classifier to estimate the category label for each drawing in the dataset.To avoid any bias due to imbalance in the distribution of drawings over categories, we randomly under sampled such that there were an equal number of drawings for each combination of geographical site (San Jose, Beijing) and the 12 categories. No additional metadata about the age of the child who produced each sketch was provided to the decoder. This procedure was repeated for each drawing in the dataset, yielding a binary a recognition score for each drawing.

@@

To assess our main hypotheses, we fit generalized linear mixed effects models to the human recognition scores to assess the factors that influenced the recognizability of the drawings that children produced. A first generalized linear mixed effect model was fit to the recognizability scores for each drawing, including fixed effects of children's age (in years), geographical site (San Jose vs.~Beijing), and drawing task (verbal cue vs.~picture cue) and the three way interaction between these key variables. We initially planned to include random slopes for the effect of drawing task on each child (as this varied within-subjects), and random slopes for the effect the full three-way interactions between task, age, and site on each category. However, models with this complicated random effects structure failed to converge, and the reported models use the maximal random effects structure that did converge -- which included random sloops for the two-way interaction between task and age on each category.

In a secondary analysis, we aimed to understand the degree to which any of the above effects were mediated by (1) children's tracing abilities and (2) the amount of effort that children expended while drawings. We thus ran the same main model while also now including fixed effects of children's estimated tracing score (see Measuring Tracing Abilities), the time children spent drawing (in seconds), the mean intensity of the drawing (i.e.~percentage of non-white pixels), and the number of strokes children used. All predictors were scaled to have a mean of 0 and a standard deviation of 1. Finally, we also assessed the degree to which tracing ability development differed by geographical site, where tracing scores were modeled as a function of age (in years), site, and their interaction, with the same random effects structure as the first model.

This analysis plan was pre-registered at \url{https://osf.io/qymjr}. While we had initially planned to use automated recognition scores as our main dependent variable, following prior work (Long et al., 2021), we found that these automated recognition scores were only modestly correlated with human recognition scores (\(r\) = 0.40, \(t\) = 22.78, p \textless{} .001); furthermore, descriptive plots of the automated recognition scores revealed a overall difference between the two drawing tasks that was not evident in the human recognition data. To be conservative, we thus fit all of our models to the human recognition data but have made all model classifications available for future research.

@@

\hypertarget{confirmatory-analyses}{%
\subsubsection{Confirmatory Analyses}\label{confirmatory-analyses}}

Overall, we found relative consistency in the developmental trajectory of children's drawings: Figure \ref{fig:main-results} shows the recognizability of children's drawings at each age as a function of the drawing task and the geographical site that they were located in (all model coefficients can be found in Table 1). We found steady changes in the recognizability of children's drawings across this age range, replicating prior work with an observational dataset collected at a free-standing kiosk (Long et al., 2021). Further, we found that these changes in recognizability did not depend on the drawing task children completed: children's drawings were equally recognizable when they were asked to \enquote{Draw this rabbit as it looks in the picture} versus they were simply asked \enquote{Can you draw a rabbit?}. While we anticipated that older children might produce more recognizable drawings when provided with a picture cue -- as do adults when presented with canonical photographs of object categories (Yang \& Fan, 2021) -- this was not the case. However, we did observe a main effect of geographical site: we found that children from Beijing, China overall tended to produce more recognizable drawings than children from San Jose, USA (see coefficients in Table 1). A closer look comparing the recognizability of the drawings made at children at each age and site hints that the children from Beijing, China may tend to produce more recognizable drawings earlier in development.
